Factor Analysis on Citations, Using a Combined Latent and Logistic Regression Model

### 2019-08-20 Manuscript for submission/manuscript.tex
Update
minor changes

\documentclass[AMS,STIX1COL]{WileyNJD-v2}
\usepackage{graphicx,times,verbatim,amsmath,amssymb,amsthm}
\usepackage{amsmath,amsfonts}
\usepackage{bbm}
\usepackage{mathtools}
\usepackage{float}  % For fixing the image in draft
\usepackage{lscape}
\usepackage[nottoc,notlot,notlof]{tocbibind}

@@

\title{Factor Analysis on Citations, using a Combined Latent and Logistic Regression Model}

@@

We propose a penalized likelihood estimation approach as follows:
\begin{eqnarray}
\label{eq:logistic10}
(\hat{\alpha}, \widehat{L}, \widehat{S})
= \mbox{arg min}_{\alpha, L, S} \left\{
-\frac{1}{n} \mathbb{L}_n(\alpha,  L, S; X) + \gamma \|S\|_1
+ \delta \|L\|_\ast\right\},
\end{eqnarray}
where $\gamma>0$ and $\delta>0$ are algorithmic parameters whose values will be discussed later,
the $L_1$ norm of matrix $S$ is defined as $\|S\|_1 = \sum_{i\neq j} S_{ij}$ (Note that we do not penalize the diagonal entries of $S$), and nuclear norm of matrix $L$ is defined as $\|L\|_\ast = {\mbox{trace}\sqrt{(L^T L)}}$.
Recall that both $S$ and $L$ are symmetric matrices.
The entries of matrix $S$ can either be positive or negative.
Note that we have imposed the diagonal entries of the matrix $X$ to be zeros.
Given that $L = F^T D F$ where matrix $D$ is diagonal with nonnegative diagonal entries, it is easy to see that matrix $L$ is positive semidefinite; which consequently leads to $\|L\|_\ast = \mbox{trace}(L)$, which is a linear functional to the matrix $L$.
The nuclear norm of $L$ penalizes the number of nonzero eigenvalues of $L$, which is the same as the rank
of $L$.
The regularization based on the nuclear norm was proposed in \cite{fazel2001rank}
and its statistical properties are studied in \cite{bach2008consistency}.

@@

\section{ Numerical analysis and Applications}
\label{sec:Num-anal}
In this section, we have conducted a small empirical study of measuring the performance of our proposed method with artificially specified graphical structures,\ref{sec:Syn-dat}.
Also we perform a real data analysis with citation network for statisticians,\ref{sec:citation_network}.
\subsection{Synthetic data}
\label{sec:Syn-dat}
\textbf{Data Generation}.
Following the notation of our paper, each edge $X_{ij}$ follows Bernoulli distribution, whose parameter is parameterized by the probability, $P^{*}_{ij}=\frac{\exp (
\alpha^{*}+F^{*T}_{i}D^{*}F^{*}_{j} +  S^{*}_{ij})}{1+\exp(\alpha^{*}+F^{*T}_{i}D^{*}F^{*}_{j} +  S^{*}_{ij})}$.
We set the model parameters randomly following these steps:
\begin{enumerate}
    \item Generate $\alpha^*$ from Unif$[-3,-2]$ (i.e., $\alpha^*\sim \text{Unif}[-3,-2]$).
    \item For the binary factor loading matrix $F^*(\in \mathbb{R}^{n \times K})$, first fill in each row of the matrix with $\floor{\frac{n}{K}}$ ones,
     and make sure that each column of $F^{*}$ has only one $1$.
     In last row of $F^*$, fill in the last remaining $n-K\floor{\frac{n}{K}}$ entries with $1$s.
     Then, we randomly choose one of columns, and fill that column with ones.
     Lastly, set $F^*=F^*J$, where $J=I_{n}-\frac{1}{n}\mathbbm{1}\mathbbm{1}^T$.
     %and project the column space of $F$ onto the orthogonal complement subspace of $\mathbbm{1}$ vector by multiplying the aforementioned $J = I_{n}-\frac{1}{n}\mathbbm{1}\mathbbm{1}^T$ to $F$.
    \item Each element of diagonal matrix $D^{*}$ are generated from Unif$[7,8]$.
    \item For the sparse component $S^*$, first draw $nnz$ distinct random pairs of indices that are uniformly distributed over the indices of upper triangular part (off diagonal) of matrix $S^*$. For each drawn pair of indices $(i,j)$, draw random number from Unif$[7,8]$ (i.e., $S^*_{ij}\sim\text{Unif}[7,8]$). $nnz$ denotes the number of non-zero component of the upper-triangular part of $S^{*}$.
\end{enumerate}

>>>>

\documentclass[AMS,STIX1COL]{WileyNJD-v2}
%\documentclass[AMS]{WileyNJD-v2}
\usepackage{graphicx,times,verbatim,amsmath,amsthm}
%\usepackage{graphicx,times,verbatim,amsmath,amssymb,amsthm}
\usepackage{amsmath,amsfonts}
\usepackage{bbm}
\usepackage{mathtools}
\usepackage{float}  % For fixing the image in draft
\usepackage{lscape}
\usepackage[nottoc,notlot,notlof]{tocbibind}

@@

\title{Factor Analysis on Citations, Using a Combined Latent and Logistic Regression Model}

@@

We propose a penalized likelihood estimation approach as follows:
\begin{eqnarray}
\label{eq:logistic10}
(\hat{\alpha}, \widehat{L}, \widehat{S})
= \mbox{arg min}_{\alpha, L, S} \left\{
-\frac{1}{n} \mathbb{L}_n(\alpha,  L, S; X) + \gamma \|S\|_1
+ \delta \|L\|_\ast\right\},
\end{eqnarray}
where $\gamma>0$ and $\delta>0$ are algorithmic parameters whose values will be discussed later,
the $L_1$ norm of matrix $S$ is defined as $\|S\|_1 = \sum_{i\neq j} S_{ij}$ (Note that we do not penalize the diagonal entries of $S$), and nuclear norm of matrix $L$ is defined as $\|L\|_\ast = {\mbox{trace}\sqrt{(L^T L)}}$.
Recall that both $S$ and $L$ are symmetric matrices.
The entries of matrix $S$ can either be positive or negative.
Note that we have imposed the diagonal entries of the matrix $X$ to be zeros.
Given that $L = F^T D F$ where matrix $D$ is diagonal with nonnegative diagonal entries, it is easy to see that matrix $L$ is positive semidefinite; which consequently leads to $\|L\|_\ast = \mbox{trace}(L)$, which is a linear functional to the matrix $L$.
The nuclear norm of $L$ mimicks the number of nonzero eigenvalues of $L$, which is the same as the rank
of $L$.
The regularization based on the nuclear norm was proposed in \cite{fazel2001rank}
and its statistical properties are studied in \cite{bach2008consistency}.

@@

\section{ Numerical analysis and Applications}
\label{sec:Num-anal}
In this section, we have conducted a small empirical study of measuring the performance of our proposed method with artificially specified graphical structures,\ref{sec:Syn-dat}.
Also we perform a real data analysis with citation network for statisticians,\ref{sec:citation_network}.
\subsection{Synthetic data}
\label{sec:Syn-dat}
\textbf{Data Generation}.
Following the notation of our paper, each edge $X_{ij}$ follows Bernoulli distribution, whose parameter is parameterized by the probability, $P^{*}_{ij}=\frac{\exp (
\alpha^{*}+F^{*T}_{i}D^{*}F^{*}_{j} +  S^{*}_{ij})}{1+\exp(\alpha^{*}+F^{*T}_{i}D^{*}F^{*}_{j} +  S^{*}_{ij})}$.
We set the model parameters randomly following these steps:
\begin{enumerate}
    \item Generate $\alpha^*$ from Unif$[-3,-2]$ (i.e., $\alpha^*\sim \text{Unif}[-3,-2]$).
    \item For the binary factor loading matrix $F^*(\in \mathbb{R}^{n \times K})$, first fill in each row of the matrix with $\floor{\frac{n}{K}}$ ones,
     and make sure that each column of $F^{*}$ has only one $1$.
     In last row of $F^*$, fill in the last remaining $n-K\floor{\frac{n}{K}}$ entries with $1$'s.
     Then, we randomly choose one of columns, and fill that column with ones.
     Lastly, set $F^*=F^*J$, where $J=I_{n}-\frac{1}{n}\mathbbm{1}\mathbbm{1}^T$.
     %and project the column space of $F$ onto the orthogonal complement subspace of $\mathbbm{1}$ vector by multiplying the aforementioned $J = I_{n}-\frac{1}{n}\mathbbm{1}\mathbbm{1}^T$ to $F$.
    \item Each element of diagonal matrix $D^{*}$ are generated from Unif$[7,8]$.
    \item For the sparse component $S^*$, first draw $nnz$ distinct random pairs of indices that are uniformly distributed over the indices of upper triangular part (off diagonal) of matrix $S^*$. For each drawn pair of indices $(i,j)$, draw random number from Unif$[7,8]$ (i.e., $S^*_{ij}\sim\text{Unif}[7,8]$). $nnz$ denotes the number of non-zero component of the upper-triangular part of $S^{*}$.
\end{enumerate}
